---
title: 'ECMA 31360 PSet 3'
author: "Phalgun Garimella"
fontsize: 12pt
output:
  pdf_document: default
  html_document:
    df_print: paged
  word_document: default
header-includes: \usepackage{pdfpages}
geometry: margin=0.5in
fig_crop: no
---

```{r echo=F, include = FALSE}
library(knitr)
knitr::opts_chunk$set(fig.width=4, fig.height=3, message=F, warning=F)
options(width=70, digits=6, scipen=8)
library(mosaic)
library(ggplot2)
library(readxl)
library(sandwich)
library(lmtest)
library(car)
library(ivreg)
library(systemfit)
set.seed(36)
```

# Part 1: Describe the Data

## Question 1

```{r}
control = read.csv("nswre74_control.csv")
treated = read.csv("nswre74_treated.csv")
df = rbind(control, treated)
dplyr::tally(dplyr::group_by(df, treat))
dplyr::group_by(df, treat) %>% dplyr::summarise_all(list(mean))
```

We can insert table here/on shared Overleaf document.

# Part 2: Test Balance

## Question 1

```{r}
columns = colnames(df)
opvs = columns[-c(1,10)]
p_vals = c()
formula_list = c()
null_system = c() # for Q2.2b
intercepts = c() # for Q2.2b
coefficients = c() # for Q2.2b
for (i in opvs) {
  formula = formula(paste(i, " ~ ", columns[1]))
  null_formula = formula(paste(i, " ~ ", 1))
  formula_list = c(formula_list, formula)
  null_system = c(null_system, null_formula)
  lm_model = lm(formula, data = df)
  summary_coeffs = summary(lm_model)$coefficients
  print(i)
  print(summary_coeffs)
  p_vals = c(p_vals, summary_coeffs[2,4])
  intercepts = c(intercepts, summary_coeffs[1,1])
  coefficients = c(coefficients, summary_coeffs[2,1])
}
```

In order to make conclusions regarding balance of OPVs, let us output p-values for the 10 t-tests below for convenience:

```{r}
out = t(matrix(round(p_vals, 4)))
colnames(out) = opvs
rownames(out) = "p-values"
out
```

We observe that at the 5% level, we reject the null hypothesis that the mean for `nodegree` is the same in the control and treated groups. For the 9 other OPVs, however, we fail to reject the null hypothesis that each variable's mean is the same in the control and treated groups.

## Question 2

### a

```{r}
sur_fit = systemfit::systemfit(formula_list, data = df, method = "SUR")
summary(sur_fit)
```

There does not appear to be a difference between the estimated coefficients and their SEs using the SUR system and those obtained in $\textbf{Question 1}$.

### b

```{r}
null_fit = systemfit(null_system, data = df, method = "SUR")
summary(null_fit)
lrtest_obj = lrtest(null_fit, sur_fit)
lrtest_obj
```

Let us now test joint balance manually. We will be testing the following joint hypotheses:

$H_0 : $

$H_1 : $

```{r}
n = dim(df)[1]
cov_mat = matrix(0, 10, 10)
df_opvs = df[,-c(1,10)]
for (i in 1:dim(cov_mat)[1]) {
  for (j in 1:dim(cov_mat)[2]) {
    u_i = as.matrix(df_opvs[,i] - (df[,1] * coefficients[i] + intercepts[i]))
    u_j = as.matrix(df_opvs[,j] - (df[,1] * coefficients[j] + intercepts[j]))
    cov_mat[i,j] = t(u_i) %*% u_j / (n - 2)
  }
}
cov_mat
null_cov_mat = matrix(0, 10, 10)
null_coeffs = c()
for (i in opvs) {
  null_formula = formula(paste(i, " ~ ", 1))
  null_lm_model = lm(null_formula, data = df)
  null_summary_coeffs = summary(null_lm_model)$coefficients
  null_coeffs = c(null_coeffs, null_summary_coeffs[1,1])
}
for (i in 1:dim(null_cov_mat)[1]) {
  for (j in 1:dim(null_cov_mat)[2]) {
    u_i = as.matrix(df_opvs[,i] - null_coeffs[i])
    u_j = as.matrix(df_opvs[,j] - null_coeffs[j])
    null_cov_mat[i,j] = t(u_i) %*% u_j / (n - 1)
  }
}
null_cov_mat
lr_test_stat = (n - 2) * (log(norm(null_cov_mat)) - log(norm(cov_mat)))
chi_sq_test_stat = -2 * lr_test_stat * 10
chi_sq_test_stat
pchisq(chi_sq_test_stat, df = 10, lower.tail = FALSE)
```

